ProbabilityBoundsAnalysis.jl:\\ Arithmetic with sets of distributions

### 2021-10-05 paper/paper.tex
adds minor changes

An arithmetic of probability distributions has long held interest among mathematicians and scientists. Indeed it was Kolmogorov who originally asked about the result of an operation between two random variables without knowing their joint distribution. This was answered for the sum by Makarov \cite{makarov1982estimates}, who showed the result was a set of distributions and was able to provide bounds on this function. Sklar, Schweizer and Frank generalised this result to other positive binary operations \cite{frank1987best,schweizer2011probabilistic}. In this pursuit they created copulas, a general way to encode probabilistic dependence independently from marginals. These are now an essential object used in probabilistic modelling. In his dissertation \cite{williamson1989probabilistic}, Williamson described an algorithm for efficiently performing these arithmetic operations, which can give guaranteed bounds on probability distributions in terms of an upper and lower cumulative distribution function (cdf). He called his method \textit{probabilistic arithmetic} and his sets of distributions \textit{dependency bounds}. Since then the method has been generalised \cite{ferson2015constructing,ferson1996whereof,ferson2004arithmetic} to most of the base binary and unary operations that would be present in a programming language. Probability boxes (p-boxes) are the name now given to these structures

@@

Multiple dispatch allows Julia packages with completely different functionalities to work together with minimal effort. One of the great successes of \texttt{IntervalArithmetic.jl} is that it allows for intervals to be propagated through nearly any Julia function, including those from different packages. The same paradigm allows \texttt{ProbabilityBoundsAnalysis.jl} to propagate distributions and p-boxes through general Julia functions.

@@

An uncertain programming language would require \textit{source-to-source} transformations at times. For example, one way to evaluate a programs control-flow (\texttt{if-else} conditions) with uncertain inputs is to unravel the control flow, and to evaluate each possible conditional branch uniquely, and then combine the output. For example consider following function

@@

An interval evaluation of the function $\text{unc\_flow}([-1,1])$ would be the union of the output of the different branches, the interval union $[-2, -1) \cup [0, 1]$. For a language like Python, this would require an uncertainty pre-compiler (like that proposed in \cite{gray2019computing}) whereby the python code is first transformed using a pre-compiler, and then evaluated in Python. In Julia, this can be handled natively with its metaprogramming capabilities.

@@

@@

The $\tau$ and $\rho$ convolutions can be efficiently computed in terms of the inverse representation using in \textit{ProbabilityBoundsAnalysis}:

>>>>

An arithmetic of probability distributions has long held interest among mathematicians and scientists. Indeed it was Kolmogorov who originally asked about the result of an operation between two random variables without knowing their joint distribution. This was answered for the sum by Makarov \cite{makarov1982estimates}, who showed the result was a set of distributions and was able to provide bounds on this function. Sklar, Schweizer and Frank generalised this result to other positive binary operations \cite{frank1987best,schweizer2011probabilistic}. In this pursuit they created copulas, a general way to encode probabilistic dependence independently from marginals. These are now an essential object used in probabilistic modelling. In his dissertation \cite{williamson1989probabilistic}, Williamson described an algorithm for efficiently performing these arithmetic operations, which can give guaranteed bounds on probability distributions in terms of an upper and lower cumulative distribution function (cdf). He called his method \textit{probabilistic arithmetic} and his sets of distributions \textit{dependency bounds}. Since then the method has been generalised \cite{ferson2015constructing,ferson1996whereof,ferson2004arithmetic} to most of the base binary and unary operations that would be present in a programming language. Probability boxes (p-boxes) are the name now given to these structures.

@@

Multiple dispatch allows Julia packages with completely different functionalities to work together often little extra effort. One of the great successes of \texttt{IntervalArithmetic.jl} is that it allows for intervals to be propagated through nearly any Julia function, including those from different packages. The same paradigm allows \texttt{ProbabilityBoundsAnalysis.jl} to propagate distributions and p-boxes through general Julia functions.

@@

An uncertain programming language would require \textit{source-to-source} transformations at times. For example, one way to evaluate a programs control-flow (\texttt{if-else} conditions) with uncertain inputs is to use metaprogramming. The programs control flow could be unraveled, and with each possible conditional branch evaluated for the possible instance of the variable, and then combined as an output. For example consider following function

@@

An interval evaluation of the function $\text{unc\_flow}([-1,1])$ would be the union of the output of the different branches, the interval union $[-2, -1) \cup [0, 1]$. For a language like Python, this would require an uncertainty pre-compiler (like that proposed in \cite{gray2019computing}) whereby the python code is first parsed using a separate program, and then evaluated in Python. In Julia, this can be handled natively with its metaprogramming capabilities.

@@

\iffalse
\subsubsection{Parallelism and performance} \hfill \break

Ultimately, high performant engineering and physics codes are written in

\fi

@@

The $\tau$ and $\rho$ convolutions can be efficiently computed in terms of the inverse representation using in \texttt{ProbabilityBoundsAnalysis.jl}:
